[Re] Non-additive coupling enables propagation of synchronous spiking activity in purely random networks

### 2018-03-12 article/Caze_2018.tex
Various text changes (still incomplete)

In addition to random initial conditions, i.e. random initial values of the membrane potential, the simulation includes 1--50 random spikes ``initially in transit''. The original article does not go into further details about their implementation, we decided to use the following: at the beginning of the simulation, a uniformly distributed number between 1 and 50 is drawn to determine the total number of initial spikes. These spikes are ``in transit'', meaning that they are modelled as arising from events before the start of the simulation. Since the synaptic delay in the network is \SI{5}{\milli\second}, we draw the arrival times for these spikes between \SI{0}{\milli\second} and \SI{5}{\milli\second}. Consistent with the probability for excitatory connections in the network ($p_\text{ex}=0.5$), we take half of these spikes as arising from excitatory and half from inhibitory sources, modelling there effect as described above.

@@

For the grid search presented in Fig.~\ref{fig:grid}, we varied inhibitory and excitatory synaptic weights between \SI{0.16}{\milli\volt} and \SI{0.4}{\milli\volt} (given that each neuron receives on average 150 excitatory and 150 inhibitory inputs, this corresponds to a ``total weight'' between \SI{24}{\milli\volt} and \SI{60}{\milli\volt}, as used on the axes of Fig.~\ref{fig:grid}), using 100 steps for each variable (comparable to the 96 steps used in the original publication). Each network was simulated 10 times with different initial conditions (synaptic connections and initial membrane potential values), whereas the original publication used 20 simulations per weight combination.

Our replication present slight differences with the original grid search. The stable region surface ends up earlier in a regime with strong inhibitory connections. The odd blue-yellow-red region presents in the original article differs from the continuous transition we obtained in our simulations. Implementation details might explain these differences

To numerically derive the evolution of synchronous pulse sizes presented in Fig.~\ref{fig:markov} (for the semi-analytical approach, see the next section), we performed simulations of \SI{56}{\milli\second} each. During each simulation, we stimulated the network at \SI{50}{\milli\second} with a synchronous pulse of activity, consisting of between 1 and 176 neurons (varied in steps of 5 -- the original study used group sizes between 1 and 181 varied in steps of 6). We then observed the number of active neurons at \SI{55}{\milli\second}, i.e. at one $\tau$ after the stimulation. For each stimulation group size, we ran the simulations 50~times with different initial conditions, consistent with the original study.

\subsection*{Semi-analytical and analytical methods}
We did not attempt to replicate the analytical solution for the evolution of synchronous pulses based on diffusion approximation (blue dots in Fig.~4 of \cite{Memmesheimer2012}). The original work explains in more details the semi-analytical work that was more straightforward to reproduce.
To calculate the semi-analytical solutions for Fig.~\ref{fig:markov}, based on a numerical estimate of the distribution $P(V)$, we proceeded as follows. First, we ran $2\times 50$ (once for the linear, once for the non-linear network) simulations of \SI{250}{\milli\second} each of a network with the default parameters but without any additional synchronous stimulation.  We binned all membrane potential values over all time steps and neurons into 100 bins in the voltage interval $[-\Theta/8, \Theta]$, to obtain an estimate of $P(V)$, the distribution of the membrane potential. From this distribution we can estimate the probability $F$ that an additional input $\epsilon$ drives the neuron over its threshold (or equivalently, that the neuron is less than $\epsilon$ away from its threshold):
\begin{equation}
F(\epsilon) = P(V>\Theta-\epsilon) = \int_{\Theta-\epsilon}^{\Theta}P(V)dV
\end{equation}
Let $\epsilon(n_\text{ex}, n_\text{in})$ be the total input to a neuron that receives $n_\text{ex}$ excitatory and $n_\text{in}$ inhibitory spikes, i.e. $\epsilon(n_\text{ex}, n_\text{in}) = \sigma(n_\text{ex}w_\text{ex}) + n_\text{in}w_\text{in}$. Let us assume that in the last iteration $g_i$ neurons spiked and that these neurons are refractory now, i.e. that $N-g_i$ neurons are available to spike in this iteration. To calculate the probability for a single (non-refractory) neuron $j$ to spike in response to these $g_i$ neurons, we will first calculate the probability that out of the $g_i$ neurons that spiked, exactly $n_\text{ex}$ neurons are connected with excitatory connections and $n_\text{in}$ neurons are connected with inhibitory connections to neuron $j$. There are $\binom{g_i}{n_\text{ex} + n_\text{in}}$ possibilities to chose $n_\text{ex} + n_\text{in}$ out of $g_i$ neurons, and $\binom{n_\text{ex} + n_\text{in}}{n_\text{ex}}$ possibilities to divide these neurons into groups of size $n_\text{ex}$ and $n_\text{in}$. This yields the total number of combinations $c(g_i, n_\text{ex}, n_\text{in})$:
\begin{equation}
c(g_i, n_\text{ex}, n_\text{in}) = \binom{g_i}{n_\text{ex} + n_\text{in}}\binom{n_\text{ex} + n_\text{in}}{n_\text{ex}} = \frac{g_i!}{n_\text{ex}!n_\text{in}!(g_i - n_\text{ex} - n_\text{in})!}
\end{equation}
The probability that these selected $n_\text{ex}$ neurons are actually connected via excitatory connections to neuron $j$, that the $n_\text{in}$ neurons are actually connected via inhibitory connections to neuron $j$, and that the remaining $g_i - n_\text{ex} - n_\text{in}$ neurons are \emph{not} connected to neuron $j$ is:
\begin{equation}
p(g_i, n_\text{ex}, n_\text{in}) = (p_0 p_\text{ex})^{n_\text{ex}} (p_0 p_\text{in})^{n_\text{in}} (1 - p_0)^{g_i - n_\text{ex} - n_\text{in}}
\end{equation}
Finally, we need to sum these probabilities up over all possible values for $n_\text{ex}$ and $n_\text{in}$, multiplying it with the probability that the generated input elicits a spike. This yiels the total probability $P^s(g_i)$ that a neuron $j$ spikes in response to the activity of $g_i$ neurons in the previous iteration:
\begin{equation}
P^s(g_i) = \sum_{n_\text{ex}=1}^{g_i}\sum_{n_\text{in}=0}^{g_i-n_\text{ex}} F\big(\epsilon(n_\text{ex}, n_\text{in})\big) c(g_i, n_\text{ex}, n_\text{in}) p(g_i, n_\text{ex}, n_\text{in})
\end{equation}
In Fig.~\ref{fig:markov}, we plot the expected value of spiking neurons during the next iteration, i.e.:
\begin{equation}
E(g_{i+1}|g_i = g_i') = (N - g_i') P^s(g_i')
\end{equation}

@@

To confirm the equivalence of our network's behaviour with the original study over a wide range of synaptic connection strength we ran a grid search exploration for linearly (Fig.~\ref{fig:grid}A) and non-linearly (Fig.~\ref{fig:grid}B) coupled networks. Network behaviour is colour-coded as being part of one out of four classes: stable activity, but no persistent propagation (green); stable activity with persistent propagation (blue); unstable activity after synchronous stimulation (yellow); unstable activity before synchronous stimulation (red). For quantitative definitions, see Table~\ref{tab:color}. We run each combination of connection strengths 10 times, and combine the results to yield a color in RGB notation (values between 0 and 1 for the red, green, and blue component), according to:
\begin{equation}
R = U_1 + U_2;\quad G = E + U_2;\quad B = S,
\end{equation}
where $U_1$, and $U_2$ stand for the fraction of trials displaying unstable network activity before respectively after onset of propagation, and $S$ and $E$ for the fraction of trials displaying stable background activity with respectively without persistent propagation. We obtained this scheme directly from the authors of the original study, it was only coarsely described in the original publication. Combining individual results in this way does not necessarily lead to easily interpretable colours, but using a different scheme, e.g. colouring according to the class that occurred the most often, leads to visually similar results (not shown).

Our replication does not reproduce all details of the original plots (in particular at the top of the plots), but looks very similar on a coarse scale. Importantly, it strongly supports the original paper's main conclusion that non-linearly coupled networks have a large parameter range of stable propagation (i.e., blue-coloured points), while linearly coupled networks do not. The minor differences in our replication could be due to either our use of a clock-based simulation method  (see Fig.~\ref{fig:grid_detail}), or to the slightly different initial conditions (no spikes ``in transit'').
\begin{table}
\begin{tabular}{ c  l  p{0.32\textwidth}}
  \toprule
   & Class & Definition \\
  \midrule
  $U_1$ (red) & Unstable before stimulation & $\exists t<t_\text{stim}: a_b(t) \geq 100$ \\
  $U_2$ (yellow) & Unstable after stimulation & $\exists t>t_\text{stim}: a_b(t) \geq 100$ \\
  $E$ (green) &  Stable activity, no propagation & $\forall t: a_b(t) < 100\, \wedge\, g_\text{p} < 10$\\
  $S$ (blue) & Stable activity, persistent propagation & $\forall t: a_b(t) < 100\, \wedge\, g_\text{p} \geq 10$\\
  \bottomrule
\end{tabular}
\caption{Classification of network behaviour for Fig.~\ref{fig:grid}. Here, $a_b$ refers to the background activity, the number of neurons active during a \SI{1}{\milli\second} time bin, ignoring the time steps at $t_\text{p}^k = t_\text{stim} + k\cdot\tau$, i.e. time steps that follow the synchronous stimulation by multiples of the synaptic delay. The variable $g_\text{p}$ refers to the number of successful propagations of the synchronous activity, i.e. the number of time steps $t_\text{p}^k$ where the number of active neurons exceeds the maximal background activity $a_b$.}\label{tab:color}
\end{table}

\begin{figure}
\centering
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_l_dt_100us_gs100_mean.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs100_mean.pdf}
\end{subfigure}
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs75_mean.pdf}
\end{subfigure}
\caption{\label{fig:grid}
\textbf{Qualitative network behaviour as a function of excitatory and inhibitory connection strength for linearly (A) and non-linearly (B, C) coupled networks.} For each combination, we initiated synchronous activity with 100 (A, B) or 75 (C) neurons and assessed the stability of the temporal evolution. Blue colouring indicates stable propagation of synchrony, red indicates unstable background activity before onset of propagation, yellow indicates unstable background activity after onset of propagation, and green colouring indicates the absence of propagation (see Methods for details).\\
Replication of Fig.~3a and b in \cite{Memmesheimer2012} (\url{https://doi.org/10.1371/journal.pcbi.1002384.g003})}
\end{figure}

\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_l_dt_100us.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_nl_dt_100us.pdf}
\end{subfigure}
\caption{\label{fig:markov}
\textbf{Evolution of synchronous pulses in linearly (A) and non-linearly (B) coupled networks.} To numerically estimate the probability distribution, we show occurrences of pulse-sizes $g'_1$ in response to a pulse of size $g'_0$ by grey lines; green squares represent the corresponding mean response group sizes.\\
Replication of Fig.~4 in \cite{Memmesheimer2012} (\url{https://doi.org/10.1371/journal.pcbi.1002384.g004})}
\end{figure}

Finally, we looked at the evolution of synchronous group sizes for one combination of synaptic weights ($w_\text{ex}=w_\text{in}=\SI{0.2}{\milli\volt}$) in detail (Fig.~\ref{fig:markov}). The results confirms our previous conclusion that we can faithfully replicate the network behaviour. They match both qualitatively and quantitatively, the group size consistently declines with each iteration in linearly coupled networks (Fig.~\ref{fig:markov}A), whereas it increases for intermediate group sizes and declines otherwise in non-linearly coupled networks (Fig.~\ref{fig:markov}B). While the semi-analytic solution (red dotted line) shows a good match to the numerical results (grey lines, means as green dots), it shows some quantitative differences to the semi-analytic solution presented in Fig.~4 of the original article. The semi-analytical solution in the original article consistently overestimates the size of the group in the next iteration compared to the numerical solution. In contrast, our semi-analytical solution consistently underestimates their size in linearly coupled networks (Fig.~\ref{fig:markov}A). For non-linearly coupled networks, our solution underestimates the evolution of the group size for initial group sizes up to about 115--125 and only overestimates it for higher initial group sizes.

The result of our simulation seems shifted upward compared to the original work. This difference becomes most visible at the origin of the plot. We corrected this shift (see Fig.~\ref{fig:markov_corrected}), but we still observe a small discrepency in the non-linear case. This might come from "false positive" spikes generated by the background activity (~5 spikes more in our simulation results). Yet our work strengthen the original result demonstrating a qualitative difference between linearly coupled and non-linearly coupled networks.

@@

{\sffamily \small
   \printbibliography[title=References]
%}
\clearpage
\appendix
\section*{Supplementary Figures}
\begin{figure}
\centering
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs100_mean_detail.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_50us_gs100_mean_detail.pdf}
\end{subfigure}
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_10us_gs100_mean_detail.pdf}
\end{subfigure}
\caption{\label{fig:grid_detail}
\textbf{Qualitative network behaviour as a function of excitatory and inhibitory connection strength for non-linearly coupled networks simulated with different simulation time steps.} Color code as in Fig.~\ref{fig:grid}. Results are shown for a smaller range of connection strengths than in Fig.~\ref{fig:grid}. The total inhibitory connection strength was varied between \SI{24}{\milli\volt} and \SI{34.55}{\milli\volt}, the total excitatory connection strength was varied between \SI{34.91}{\milli\volt} and \SI{45.45}{\milli\volt}, i.e. the parameter range corresponds to a square region at the top of Fig.~\ref{fig:grid}B. Simulations where performed with simulation time steps of \SI{0.1}{\milli\second} (A; the value used for all other simulations shown), \SI{0.05}{\milli\second} (B), and \SI{0.01}{\milli\second} (C).}
\end{figure}

\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_l_dt_100us_corrected.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_nl_dt_100us_corrected.pdf}
\end{subfigure}
\caption{\label{fig:markov_corrected}
\textbf{Evolution of synchronous pulses in linearly (A) and non-linearly (B) coupled networks, corrected for the background activity.} All conventions as in Fig.~\ref{fig:markov}. The numerical solutions (grey lines and green squares) have been corrected by subtracting the background activity that is expected to occur by chance at every time step, i.e. the ``false positive'' spikes that are counted as being part of the synchronous spike. Background activity was estimated from the total activity before the stimulation with the synchronous pulse and, for a simulation time step of \SI{0.1}{\milli\second}, amounted to about 5.5 spikes per time step.}
\end{figure}
\end{document}

>>>>

In addition to random initial conditions, i.e. random initial values of the membrane potential, the simulation includes 1--50 random spikes ``initially in transit''. The original article does not go into further details about their implementation, so we decided to use the following: at the beginning of the simulation, a uniformly distributed number between 1 and 50 is drawn to determine the total number of initial spikes. These spikes are ``in transit'', meaning that they are modelled as arising from events before the start of the simulation. Since the synaptic delay in the network is \SI{5}{\milli\second}, we draw the arrival times for these spikes between \SI{0}{\milli\second} and \SI{5}{\milli\second}. Consistent with the probability for excitatory connections in the network ($p_\text{ex}=0.5$), we take half of these spikes as arising from excitatory and half from inhibitory sources, modelling their effect as described above.

@@

For the grid search presented in Fig.~\ref{fig:grid}, we varied inhibitory and excitatory synaptic weights between \SI{0.16}{\milli\volt} and \SI{0.4}{\milli\volt}. Given that each neuron receives on average 150 excitatory and 150 inhibitory inputs, this corresponds to a ``total weight'' between \SI{24}{\milli\volt} and \SI{60}{\milli\volt}, as used on the axes of Fig.~\ref{fig:grid}). We used a resolution of 100 steps for each variable, comparable to the 96 steps used in the original publication. Each network was simulated 10 times with different initial conditions (synaptic connections and initial membrane potential values), whereas the original publication used 20 simulations per weight combination.

To numerically derive the evolution of synchronous pulse sizes presented in Fig.~\ref{fig:markov} (for the semi-analytical approach, see the next section), we performed simulations of \SI{56}{\milli\second} each. During each simulation, we stimulated the network at \SI{50}{\milli\second} with a synchronous pulse of activity, consisting of between 1 and 181 neurons (varied in steps of 5 -- the original study used group sizes in the same range varied in steps of 6). We then observed the number of active neurons at \SI{55}{\milli\second}, i.e. at one $\tau$ after the stimulation. For each stimulation group size, we ran the simulations 50~times with different initial conditions, consistent with the original study.

\subsection*{Semi-analytical and analytical methods}
We did not attempt to replicate the analytical solution for the evolution of synchronous pulses based on diffusion approximation (blue dots in Fig.~4 of \cite{Memmesheimer2012}), since the original publication did not give a detailed description of this approach. However, we replicate the semi-analytical work that was described in more detail.

To calculate the semi-analytical solutions for Fig.~\ref{fig:markov}, based on a numerical estimate of the distribution $P(V)$, we proceeded as follows. First, we ran $2\times 50$ (once for the linear, once for the non-linear network) simulations of \SI{250}{\milli\second}. Each of these simulations was randomly connected and had random initial conditions, but received no additional synchronous stimulation. The original study additionally simulated each of the 100 networks 10 times, with the same synaptic connectivity but differing in the initial conditions. We then binned all membrane potential values over all time steps and neurons into 100 bins in the voltage interval $[-\Theta/8, \Theta]$ to obtain an estimate of $P(V)$, the distribution of the membrane potential. Note that the original study did not further specify how $P(V)$ was estimated from the simulation results.

We then calculate the expected synchronous group size after one iteration as in the original publication. To make this calculation clearer, we now describe its derivation in detail. From the distribution $P(V)$ we can estimate the probability $F$ that an additional input $\epsilon$ drives the neuron over its threshold (or equivalently, that the neuron is less than $\epsilon$ away from its threshold):
\begin{equation}
F(\epsilon) = P(V>\Theta-\epsilon) = \int_{\Theta-\epsilon}^{\Theta}P(V)dV
\end{equation}
Let $\epsilon(n_\text{ex}, n_\text{in})$ be the total input to a neuron that receives $n_\text{ex}$ excitatory and $n_\text{in}$ inhibitory spikes, i.e. $\epsilon(n_\text{ex}, n_\text{in}) = \sigma(n_\text{ex}w_\text{ex}) + n_\text{in}w_\text{in}$. Let us assume that in the previous iteration $g_i$ neurons spiked and that these neurons are refractory now, i.e. that $N-g_i$ neurons are available to spike in this iteration. To calculate the probability for a single (non-refractory) neuron $j$ to spike in response to these $g_i$ neurons, we will first calculate the probability that out of the $g_i$ neurons that spiked, exactly $n_\text{ex}$ neurons are connected with excitatory connections and $n_\text{in}$ neurons are connected with inhibitory connections to neuron $j$. There are $\binom{g_i}{n_\text{ex} + n_\text{in}}$ possibilities to chose $n_\text{ex} + n_\text{in}$ out of $g_i$ neurons, and $\binom{n_\text{ex} + n_\text{in}}{n_\text{ex}}$ possibilities to divide these neurons into groups of size $n_\text{ex}$ and $n_\text{in}$. This yields the total number of combinations $c(g_i, n_\text{ex}, n_\text{in})$:
\begin{equation}
c(g_i, n_\text{ex}, n_\text{in}) = \binom{g_i}{n_\text{ex} + n_\text{in}}\binom{n_\text{ex} + n_\text{in}}{n_\text{ex}} = \frac{g_i!}{n_\text{ex}!n_\text{in}!(g_i - n_\text{ex} - n_\text{in})!}
\end{equation}
The probability that these selected $n_\text{ex}$ neurons are actually connected via excitatory connections to neuron $j$, that the $n_\text{in}$ neurons are actually connected via inhibitory connections to neuron $j$, and that the remaining $g_i - n_\text{ex} - n_\text{in}$ neurons are \emph{not} connected to neuron $j$ is:
\begin{equation}
p(g_i, n_\text{ex}, n_\text{in}) = (p_0 p_\text{ex})^{n_\text{ex}} (p_0 p_\text{in})^{n_\text{in}} (1 - p_0)^{g_i - n_\text{ex} - n_\text{in}}
\end{equation}
Finally, we need to sum these probabilities up over all possible values for $n_\text{ex}$ and $n_\text{in}$, multiplying it with the probability that the generated input elicits a spike. This yiels the total probability $P^s(g_i)$ that a neuron $j$ spikes in response to the activity of $g_i$ neurons in the previous iteration:
\begin{equation}
P^s(g_i) = \sum_{n_\text{ex}=1}^{g_i}\sum_{n_\text{in}=0}^{g_i-n_\text{ex}} F\big(\epsilon(n_\text{ex}, n_\text{in})\big) c(g_i, n_\text{ex}, n_\text{in}) p(g_i, n_\text{ex}, n_\text{in})
\end{equation}
In Fig.~\ref{fig:markov}, we plot the expected value of spiking neurons during the next iteration, given a certain initial synchronous pulse $g_i'$:
\begin{equation}
E(g_{i+1}|g_i = g_i') = (N - g_i') P^s(g_i')
\end{equation}

@@

To confirm the equivalence of our network's behaviour with the original study over a wide range of synaptic connection strength we ran a grid search exploration for linearly (Fig.~\ref{fig:grid}A) and non-linearly (Fig.~\ref{fig:grid}B) coupled networks. Network behaviour is colour-coded as being part of one out of four classes: stable activity, but no persistent propagation (green); stable activity with persistent propagation (blue); unstable activity after synchronous stimulation (yellow); unstable activity before synchronous stimulation (red). For quantitative definitions, see Table~\ref{tab:color}. We run each combination of connection strengths 10 times, and combine the results to yield a color in RGB notation (values between 0 and 1 for the red, green, and blue component), according to:
\begin{equation}
R = U_1 + U_2;\quad G = E + U_2;\quad B = S,
\end{equation}
where $U_1$, and $U_2$ stand for the fraction of trials displaying unstable network activity before respectively after onset of propagation, and $S$ and $E$ for the fraction of trials displaying stable background activity with respectively without persistent propagation. We obtained this scheme directly from the authors of the original study, it was only coarsely described in the original publication. Combining individual results in this way does not necessarily lead to easily interpretable colours, but using a different scheme, e.g. colouring according to the class that occurred the most often, leads to visually similar results (Fig.~\ref{fig:grid_max}).

\begin{table}
\begin{tabular}{ c  l  p{0.32\textwidth}}
  \toprule
   & Class & Definition \\
  \midrule
  $U_1$ (red) & Unstable before stimulation & $\exists t<t_\text{stim}: a_b(t) \geq 100$ \\
  $U_2$ (yellow) & Unstable after stimulation & $\exists t>t_\text{stim}: a_b(t) \geq 100$ \\
  $E$ (green) &  Stable activity, no propagation & $\forall t: a_b(t) < 100\, \wedge\, g_\text{p} < 10$\\
  $S$ (blue) & Stable activity, persistent propagation & $\forall t: a_b(t) < 100\, \wedge\, g_\text{p} \geq 10$\\
  \bottomrule
\end{tabular}
\caption{Classification of network behaviour for Fig.~\ref{fig:grid}. Here, $a_b$ refers to the background activity, the number of neurons active during a \SI{1}{\milli\second} time bin, ignoring the time steps at $t_\text{p}^k = t_\text{stim} + k\cdot\tau$, i.e. time steps that follow the synchronous stimulation by multiples of the synaptic delay. The variable $g_\text{p}$ refers to the number of successful propagations of the synchronous activity, i.e. the number of time steps $t_\text{p}^k$ where the number of active neurons exceeds the maximal background activity $a_b$.}\label{tab:color}
\end{table}

Our replication does not reproduce all the quantitative details of the original plots. The parameter range where simulations are only unstable after they receive the synchronous stimulation (yellow) is smaller than in the original article. For linear networks, we do observe almost no simulations classified in this category, while the original article has a thin line of such classifications between the stable regime without propagation (green) and the regime where simulations are unstable even without stimulation (red). Similarly, in the original article the parameter exploration shows for each inhibitory connection weight a range of excitatory connection weights of about \SI{5}{\milli\volt} where the network is unstable only after the synchronous stimulation (yellow). In contrast, our simulations show a much smaller zone when the inhibitory connection strength is small (Fig.~\ref{fig:grid}B). It seems that these differences mostly stem from the fact that the transition to the regime where simulations are unstable without stimulation (red) happens for lower excitatory and higher inhibitory connection strength in our simulations than in the original article, i.e. that the red zone is shifted towards the bottom left. This is particularly evident for strong excitatory and inhibitory connections in non-linearly coupled networks (bottom right in Fig.~\ref{fig:grid}B and C) which are unstable in our simulations, but stable or only unstable after stimulation in the original publication. Finally, non-linearly coupled networks show a discontinuity for weak inhibitory connection strengths in the original publication. For increasing excitatory connection strengths, network behaviour goes from stable (blue) to unstable after stimulation (yellow) to unstable before stimulation (red), but then get classified again as stable (blue) in a small range of parameters before finally transitioning to unstable activity (yellow and red). We do not observe this in our simulations.

Despite these differences, the qualitative network behaviour looks very similar on a coarse scale. Importantly, it strongly supports the original paper's main conclusion that non-linearly coupled networks have a large parameter range of stable propagation (i.e., blue-coloured points), while linearly coupled networks do not. The differences in our replication are most likely due to our use of a clock-based simulation method  (see Fig.~\ref{fig:grid_detail}). In principle, another reason could be implementation details of the initial ``spikes in transit'' that were not fully specified in the original article, but this seems unlikely as completely removing them did not change the results in a noticeable way (not shown).

\begin{figure}
\centering
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_l_dt_100us_gs100_mean.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs100_mean.pdf}
\end{subfigure}
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs75_mean.pdf}
\end{subfigure}
\caption{\label{fig:grid}
\textbf{Qualitative network behaviour as a function of excitatory and inhibitory connection strength for linearly (A) and non-linearly (B, C) coupled networks.} For each combination, we initiated synchronous activity with 100 (A, B) or 75 (C) neurons and assessed the stability of the temporal evolution. Blue colouring indicates stable propagation of synchrony, red indicates unstable background activity before onset of propagation, yellow indicates unstable background activity after onset of propagation, and green colouring indicates the absence of propagation (see Methods for details).\\
Replication of Fig.~3 in \cite{Memmesheimer2012} (\url{https://doi.org/10.1371/journal.pcbi.1002384.g003})}
\end{figure}

\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_l_dt_100us_corrected.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_nl_dt_100us_corrected.pdf}
\end{subfigure}
\caption{\label{fig:markov_corrected}
\textbf{Evolution of synchronous pulses in linearly (A) and non-linearly (B) coupled networks.} To numerically estimate the probability distribution, we show occurrences of pulse-sizes $g'_1$ in response to a pulse of size $g'_0$ by grey lines; green squares represent the corresponding mean response group sizes.\\
Replication of Fig.~4 in \cite{Memmesheimer2012} (\url{https://doi.org/10.1371/journal.pcbi.1002384.g004})}
\end{figure}

Finally, we looked at the evolution of synchronous group sizes for one combination of synaptic weights ($w_\text{ex}=w_\text{in}=\SI{0.2}{\milli\volt}$) in detail (Fig.~\ref{fig:markov}). The results confirms our previous conclusion that we can faithfully replicate the network behaviour which shows identical qualitative behaviour; the group size consistently declines with each iteration in linearly coupled networks (Fig.~\ref{fig:markov}A), whereas it increases for intermediate group sizes and declines otherwise in non-linearly coupled networks (Fig.~\ref{fig:markov}B). Quantitatively, the semi-analytical results also match well. Based on the plotted values in the original publication, the curve based on the semi-analytical solution crosses the identity line at about 85 ($G_1$) and 135 ($G_2$) for non-linearly coupled networks, peaking in between for $g'0\approx125$ and $g'1\approx139$. In our simulations, it crosses the line at about 86 and 134, and peaks at $g'0\approx126$ and $g'1\approx137$. As in the original article, our numerical results show slightly lower values for the group size in the next iteration, with the biggest difference for values around the peak of the function for non-linearly coupled networks (Fig.~\ref{fig:markov_corrected}B). However, we see a quantitative difference for low initial group sizes. In these simulations, the original publication shows a close match between the numerical results and the semi-analytical solution. In our simulation, we see %TODO

To obtain this good quantitative match, we however add to apply a correction to the numerical results (for the uncorrected values, see Fig.~\ref{fig:markov}). The activity of the full network in the absence of stimulation is about \SI{55}{\kilo\hertz} (see Fig.~\ref{fig:raster}). Given that our simulation uses a time step of \SI{0.1}{\milli\second}, this means that at every time step about 5.5 neurons on average are synchronously active. When we stimulate a network with a pulse of $g'_0$ synchronous neurons, the actual number of synchronously active neurons is therefore higher than this stimulation. Similarly, we have to correct the size of the observed group in the next iteration. In contrast to the simulation strategy of the original publication that could make use of an ``infinite precision'' to detect synchronous spikes triggered by the stimulation, our simulation will also detect all the synchronous spikes resulting from the background activity. In the corrected values shown in Fig.~\ref{fig:markov_corrected} we therefore used the actually observed synchronous spikes $\hat{g}'_0 \geq g'_0$ at the time of the stimulation, and the number of added spikes due to the stimulation $\hat{g}'_1 = g'_{1 \text{stim}} - g'_{1 \text{no stim}}$. Note that the latter correction slightly ``overcorrects'', as it discards all spikes that are resulting from background activity, even those that also would have been triggered by the stimulation.

Despite the small quantitative differences, our work strengthens the original result demonstrating a qualitative difference between linearly coupled and non-linearly coupled networks.

@@

{\sffamily \small
   \printbibliography[title=References]
%}
\clearpage
\appendix
\section*{Supplementary Figures}
\begin{figure}[h]
\centering
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_l_dt_100us_gs100.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs100.pdf}
\end{subfigure}
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs75.pdf}
\end{subfigure}
\caption{\label{fig:grid_max}
\textbf{Qualitative network behaviour as a function of excitatory and inhibitory connection strength for linearly (A) and non-linearly (B, C) coupled networks.} All conventions as in Fig.~\ref{fig:grid}, but points are coloured according the observed qualitative behaviour that occurred the most often among the 10 independent simulations for given connection strengths.}
\end{figure}

\begin{figure}[h]
\centering
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs100_mean_detail.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_50us_gs100_mean_detail.pdf}
\end{subfigure}
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_10us_gs100_mean_detail.pdf}
\end{subfigure}
\caption{\label{fig:grid_detail}
\textbf{Qualitative network behaviour as a function of excitatory and inhibitory connection strength for non-linearly coupled networks simulated with different simulation time steps.} Color code as in Fig.~\ref{fig:grid}. Results are shown for a smaller range of connection strengths than in Fig.~\ref{fig:grid}. The total inhibitory connection strength was varied between \SI{24}{\milli\volt} and \SI{34.55}{\milli\volt}, the total excitatory connection strength was varied between \SI{34.91}{\milli\volt} and \SI{45.45}{\milli\volt}, i.e. the parameter range corresponds to a square region at the top of Fig.~\ref{fig:grid}B. Simulations where performed with simulation time steps of \SI{0.1}{\milli\second} (A; the value used for all other simulations shown), \SI{0.05}{\milli\second} (B), and \SI{0.01}{\milli\second} (C).}
\end{figure}

\begin{figure}[h]
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_l_dt_100us.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_nl_dt_100us.pdf}
\end{subfigure}
\caption{\label{fig:markov}
\textbf{Evolution of synchronous pulses in linearly (A) and non-linearly (B) coupled networks, corrected for the background activity.} All conventions as in Fig.~\ref{fig:markov}. Without correction (see text)}
\end{figure}
\end{document}
